---
title: "Communicating Data Science Results"
author: "Eric Farng"
date: "November 26, 2015"
output: html_document
---

For this assignment, I looked at the crime incident data from San Francisco. Below is a histogram of the time occurance of the incident. we can see that the most incidents occur around 8PM and the fewest occur at 4AM.

```{r, echo=FALSE, fig.width=7, fig.height=7}
sanfran <- read.csv("sanfrancisco_incidents_summer_2014.csv", as.is=TRUE)
sanfran$Time_num <- sapply(strsplit(sanfran$Time,":"),
   function(x) {
   	x <- as.numeric(x)
   	x[1]+x[2]/60
   }
)


hist(sanfran$Time_num, breaks=24, xlim=c(0, 25), 
	 xlab="Hour of the Day", main="Incidents of Crime by time of day")

```

Next, I looked at the trend of different categories of crimes across the time of day. The Time of day was bucketed into 6 4-hour time periods.

```{r, echo=FALSE, fig.width=7, fig.height=7}
sanfran$Time_cat <- trunc(sanfran$Time_num / 4)
t1 <- table(sanfran[sanfran$Category=='LARCENY/THEFT',]$Time_cat)/sum(sanfran$Category=='LARCENY/THEFT')
t2 <- table(sanfran[sanfran$Category=='OTHER OFFENSES',]$Time_cat)/sum(sanfran$Category=='OTHER OFFENSES')
t3 <- table(sanfran[sanfran$Category=='NON-CRIMINAL',]$Time_cat)/sum(sanfran$Category=='NON-CRIMINAL')
t4 <- table(sanfran[sanfran$Category=='ASSAULT',]$Time_cat)/sum(sanfran$Category=='ASSAULT')
t5 <- table(sanfran[sanfran$Category=='VEHICLE THEFT',]$Time_cat)/sum(sanfran$Category=='VEHICLE THEFT')
t6 <- table(sanfran[sanfran$Category=='WARRANTS',]$Time_cat)/sum(sanfran$Category=='WARRANTS')
t7 <- table(sanfran[sanfran$Category=='DRUG/NARCOTIC',]$Time_cat)/sum(sanfran$Category=='DRUG/NARCOTIC')
t8 <- table(sanfran[sanfran$Category=='SUSPICIOUS OCC',]$Time_cat)/sum(sanfran$Category=='SUSPICIOUS OCC')
t9 <- table(sanfran[sanfran$Category=='MISSING PERSON',]$Time_cat)/sum(sanfran$Category=='MISSING PERSON')

max <- 24 / 4 - 1
plot(0:max, t1, type="l", col=2, axes=FALSE,
	 ylab="Percent of Category during Time", ylim=c(0,0.35),
	 xlab="Hour of the Day", main="Percent of Crime by Time of Day")
axis(2, at=(0:7)/100*5, tick=TRUE)
axis(1, at=0:max, tick=TRUE, labels=c("12-4AM", "4-8AM", "8-12PM", "12-4PM", "4-8PM", "8PM-12AM"))
abline(h=(0:7)/100*5, lty=3, col=1)
lines(0:max, t2, type="l", col=3, ylab="")
lines(0:max, t3, type="l", col=4, ylab="")
lines(0:max, t4, type="l", col=5, ylab="")
lines(0:max, t5, type="l", col=6, ylab="")
lines(0:max, t6, type="l", col=7, ylab="")
lines(0:max, t7, type="l", col=2, ylab="", lty=2)
lines(0:max, t8, type="l", col=3, ylab="", lty=2)
lines(0:max, t9, type="l", col=4, ylab="", lty=2)

legend("topleft", 
	   c("Larceny/Theft", "Other", "Non-criminal", "Assault", "Vehicle Theft", 
	     "Warrants", "Drug/Narcotic", "Suspicious Occ", "Missing Person"),
	   fill=c(2, 3, 4, 5, 6, 7, 2, 3, 4),
	   lty=c(1, 1, 1, 1, 1, 1, 2, 2, 2)
)
	  
```

I choose the top 9 most common crime categories. The most common is on the top of the legend and the least common is on the bottom. Since each category has a different amount of total indicents, each curve was normalized to a probability so each value is the percent of crimes during those hours. So 31% of Vehicle Theft occurred between 8PM and 12AM

We can see a general trend of crime mostly occuring later in the evening when it is dark. And the least crime happens from 4am-8am. Even criminals need to sleep.

We see that Assults are most common in the early hours (12-4AM) and they were third most common from 8PM-12AM. 

Missing Persons are most common from 8am-12PM which would be when they are mostly commonly reported.

Vehicle theft and Larceny theft occur late in the evening. Non-criminal crime happens mostly during the day.

Below are the number of incidents for each time. Larceny/Theft have the most incidents throughout the day. We can see that Missing persons have a higher portion of the incidents from 8AM to Noon, but still has fewer incidents than Larceny/Theft


```{r, echo=FALSE, fig.width=7, fig.height=7}
sanfran_common <- sanfran[sanfran$Category %in% c("LARCENY/THEFT", "OTHER OFFENSES", "NON-CRIMINAL", "ASSAULT", "VEHICLE THEFT", "WARRANTS", "DRUG/NARCOTIC", "SUSPICIOUS OCC", "MISSING PERSON"),]

t1 <- table(sanfran_common[sanfran_common$Time_cat==5,]$Category)
pie(t1, labels=names(t1), main="Number of Indencents from 8PM-Midnight")

t1 <- table(sanfran_common[sanfran_common$Time_cat==2,]$Category)
pie(t1, labels=names(t1), main="Number of Indencents from 8AM-Noon")

```

